---
title: "Week 7 Data Exploration Project"
author: "Chequala Fuller"
date: "2/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Before reading in the necessary data, we will need to load all libraries required to execute this analysis. 

```{r, echo = FALSE}
source("../code/01_dep_load_libraries.R")
```

## Read files

Once libraries are successfully loaded, read in the appropriate data files. For this data exploration project, we will be using the provided _Most+Recent+Cohorts+(Scorecard+Elements).csv_, _id_name_link.csv_, and twelve data files with the naming convention pattern _trend_up_to_xx.csv_ to perform our analysis.  To read these data files, we use the _list.files()_ and _map()_ functions from the _tidyverse_ package as well as the _read_csv_ from the built-in _readr_ package. The following dataframes, values, and functions have been assigned and/or created to prepare us for tidy data.


Dataframes         | No. Observations | No. Variables
------------------ | ---------------- | -------------
recent_cohorts     | 7804             | 122
id_name_link       | 3555             | 3
compile_data       | 1593740          | 6

Values             | Type
------------------ | ----------------
file_list          | chr [1:12]

Functions          | Parameter(s)
------------------ | ----------------
process_file       | function(filename)


```{r, echo = FALSE, results = 'hide'}
source("../code/02_dep_read_data.R")
```


## Tidy Data

Now, that we have all of our desired data read, let's clean it up and prepare it for the regression model(s), graph(s) and analysis.


```{r, echo = FALSE, results = 'hide'}
source("../code/03_dep_tidy_data.R")
```



## Data Analysis


```{r}
source("../code/04_dep_data_analysis.R")
```


## The Writeup

The College Scorecard was released at the start of September 2015. Among colleges that predominantly grant bachelor’s degrees, did it result in more student interest in high-earnings colleges relative to low-earnings ones (as proxied by Google searches for keywords associated with those colleges)?

  a. how can we define “high-earning” and “low-earning” colleges? There’s         not a single answer - be ready to defend your choice.
  
  b. What level should the data be at?
  
  c. How should the regression model be designed to answer the question           (transformations and functional form? Standard error adjustments?          etc.), and how can we interpret the results once we have them?


